---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Following piece of code extracts the zip file in data directory and reads it into activityData variable.
```{r}
#system("unzip activity.zip -d data/")
activityData <- read.csv("data/activity.csv")
library(dplyr)
library(ggplot2)
```


## What is mean total number of steps taken per day?

```{r results='asis'}
completeData <- subset(activityData, complete.cases(activityData))
groupedData <- completeData %>% group_by(date) %>% summarise(steps = mean(steps))
g <- ggplot(groupedData, aes(date)) + geom_histogram(aes(y = steps), stat = "identity")
g <- g + theme(axis.text.x = element_text(angle=90))
print(g)

averageSteps <- mean(groupedData$steps)
medianSteps <- median(groupedData$steps)
```

The mean of total number of steps per day is `r averageSteps` and median is `r medianSteps`

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
