---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Following piece of code extracts the zip file in data directory and reads it into activityData variable.
```{r message=FALSE}
#system("unzip activity.zip -d data/")
activityData <- read.csv("data/activity.csv")
library(dplyr)
library(ggplot2)
dim(activityData)
```


## What is mean total number of steps taken per day?

```{r message=FALSE,results='asis'}
completeData <- subset(activityData, complete.cases(activityData))
dataByDate <- completeData %>% group_by(date) %>% summarise(steps = sum(steps))
averageSteps <- round(mean(dataByDate$steps), digits = 0)
medianSteps <- median(dataByDate$steps)
g <- ggplot(dataByDate, aes(steps)) + geom_histogram(binwidth = 2000) +
  geom_vline(aes(xintercept=averageSteps), color="red", linetype="dashed", size=1) +
  xlab("Steps Per Day") + ylab("Frequency") + ggtitle("Histogram: Steps Per Day")
print(g)
options(scipen=999)
```

Note: This is a historgram as opposed to a bar plot clearly showing frequency of distribution.

The mean of total number of steps per day is `r averageSteps` and median is `r medianSteps`

## What is the average daily activity pattern?
```{r message=FALSE,results='asis'}
dataByInterval <- completeData %>% group_by(interval) %>% summarise(steps = mean(steps))
maxPoint <- data.frame(dataByInterval[dataByInterval$steps == max(dataByInterval$steps),])
g <- ggplot(dataByInterval, aes(interval, steps)) + geom_line() +
  xlab("Interval") + ylab("Average Steps") +
  geom_point(data = maxPoint, aes(interval, steps), color = "red") +
  geom_text(data = maxPoint, aes(x = interval * 1.15, y = steps, label = "Max", color = "red")) +
  theme(legend.position = "none")
    
print(g)
```

As the above graph shows, 5-minute interval with maximum steps is `r maxPoint[1]`

## Imputing missing values
```{r warning=FALSE,message=FALSE,results='asis'}
missingData <- activityData[!complete.cases(activityData),]
numMissing <- length(missingData)
source("FillIn.R")
imputedData <- FillIn(activityData, dataByInterval, "steps", "steps", c("interval"))
imputedDataByDate <- imputedData %>% group_by(date) %>% summarise(steps = sum(steps))
imputedAverageSteps <- round(mean(imputedDataByDate$steps), digits = 0)
imputedMedianSteps <- median(imputedDataByDate$steps)
g <- ggplot(imputedDataByDate, aes(steps)) + geom_histogram(binwidth = 2000) +
  geom_vline(aes(xintercept=imputedAverageSteps), color="red", linetype="dashed", size=1) +
  xlab("Steps Per Day") + ylab("Frequency") + ggtitle("Histogram: Steps Per Day")
print(g)
```

Total number of missing values: `r dim(missingData)[1]`

The mean of total number of steps per day is `r imputedAverageSteps` and median is `r imputedMedianSteps`

The impact of imputing missing data is not visible on mean value given that mean calculated per interval is used to impute missing values. Median did change towards a slightly lower value.

## Are there differences in activity patterns between weekdays and weekends?
```{r message=FALSE,results='asis'}
library(lubridate)
imputedData$day <- as.factor(ifelse(weekdays(ymd(imputedData$date)) %in% c("Saturday","Sunday"), "Weekend", "Weekday"))
groupByDayInterval <- imputedData %>% group_by(day, interval) %>% summarise(steps = mean(steps))
g <- ggplot(groupByDayInterval, aes(x = interval, y = steps, color = day)) + geom_line() + facet_wrap(~day, nrow = 2) + 
  xlab("Interval") + ylab("Average Steps") + ggtitle("Comparison of Average Steps (Weekday vs Weekend)")
print(g)
```

One major thing visible is during weekdays there is peak around interval 800, while weekend seems to have a lot more activity throughout the day. Another way to see the difference is by plotting Weekday and Weekend on the same plot.

```{r message=FALSE,results='asis'}
groupByDayInterval <- imputedData %>% group_by(day, interval) %>% summarise(steps = mean(steps))
g <- ggplot(groupByDayInterval, aes(x = interval, y = steps, color = day)) + geom_line() +
  xlab("Interval") + ylab("Average Steps") + ggtitle("Comparison of Average Steps (Weekday vs Weekend)")
print(g)
```
